---
title: "P8105 Data Science I - Homework 3"
author: "Kevin S.W."
date: "10/3/2019"
output: github_document
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

Packages <- c("tidyverse", "dplyr", "ggridges", "hexbin", "patchwork")
invisible(lapply(Packages, library, character.only = TRUE))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))

```

# Introduction
Homework for showing capability in data visualization and exploratory data analysis

# Problem 1

```{r instacart_data}

# Loads instacart data from source
library(p8105.datasets)
data("instacart")

# Assigns a new dataframe to the data-source to prevent frequent "reloads"
### Not sure if necessary.
insta_df <- instacart %>%  
  janitor::clean_names()

```

The InstaCart dataset collects data on customers and the things they order from local groceries. Data that we're using are cleaned previously and is limited because it only contains data from `eval_set = train`; likely dataset used in model-building. The dataframe contains `r nrow(instacart)` observations from `r ncol(instacart)` categories including `eval_set`. All these variables are linked in some way. For example, these variables are likely of particular importance:

* `order_id`: order identifier; if a person starts an order, all the items inside will have the same `order_id`
* `product_id`: product identifier which links a value to certain products (e.g. 340 = cake, 255 = soda, etc)
* `user_id`: customer identifier; unique id for customers (Peanut's id = 1000, John's id = 13,...)
* `order_dow`: the day of the week on which the order was placed (Mon, Tues, Wed,...)
* `order_hour_of_day`: the hour of the day on which the order was placed (13:00, 21:00,...)
* `product_name`: name of product such as "Organic Garlic", "Spring Water", or "Xpolsive Pizza Baked Snack Cracker"
* `aisle_id`: aisle identifier that is linked to `aisle` as a unique identifier
* `department_id`: department identifier
* `aisle`: the name of the aisle such as fresh vegetables or yogurt
* `department`: the name of the department (canned goods, snacks, frozen, ...)

```{r instacart_explore}

aisle_summary <- insta_df %>% 
  select(product_name, aisle_id, aisle) %>% 
  drop_na() %>% 
  mutate(
    aisle = as.factor(aisle),
    aisle_id = as.factor(aisle_id)
  ) %>% 
  group_by(aisle, aisle_id) %>% 
  summarize(
    aisle_order_count = n()
  ) %>% 
  arrange(desc(aisle_order_count))

```

As indicated above, this is a very large dataset. It contains `r nrow(aisle_summary)` unique aisles and `r pull(aisle_summary, aisle)[pull(aisle_summary, aisle_order_count) == max(pull(aisle_summary, aisle_order_count))]` is/are the most ordered item with `r max(pull(aisle_summary, aisle_order_count))` orders. A graph below illustrates the spread of the number of items ordered in each aisle for items with more than 10,000 items ordered. 

```{r }

aisle_summary %>% 
  filter(aisle_order_count > 10000) %>% 
  ggplot(aes(x = aisle_order_count, color = aisle)) +
    geom_bar(alpha = 0.3, binwidth = 5) +
    #labs(title = "***", 
     #  subtitle = "***",
      # x = "***",
       #y = "***",
      # color = "***") +
  theme(plot.title = element_text(hjust = 0.5, size = 18),
        plot.subtitle = element_text(hjust = 0.5, size = 8),
        legend.position = "none")

```


* Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

* Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

* Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

# Problem 2

For this problem, we will be loading the BRFSS data included in the p8105 package.

```{r brfss_load}

brfss_df <- data("brfss_smart2010")
  

```


First, do some data cleaning:

format the data to use appropriate variable names;
focus on the “Overall Health” topic
include only responses from “Excellent” to “Poor”
organize responses as a factor taking levels ordered from “Poor” to “Excellent”
Using this dataset, do or answer the following (commenting on the results of each):

In 2002, which states were observed at 7 or more locations? What about in 2010?
Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).
Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.
Problem 3
Accelerometers have become an appealing alternative to self-report techniques for studying physical activity in observational studies and clinical trials, largely because of their relative objectivity. During observation periods, the devices measure “activity counts” in a short period; one-minute intervals are common. Because accelerometers can be worn comfortably and unobtrusively, they produce around-the-clock observations.

This problem uses five weeks of accelerometer data collected on a 63 year-old male with BMI 25, who was admitted to the Advanced Cardiac Care Center of Columbia University Medical Center and diagnosed with congestive heart failure (CHF). The data can be downloaded here. In this spreadsheet, variables activity.* are the activity counts for each minute of a 24-hour day starting at midnight.

Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).
Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?
Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.
Optional post-assignment survey
If you’d like, a you can complete this short survey after you’ve finished the assignment.
